---
title: "Getting started with forecastsuite"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with forecastsuite}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4)
```

forecastsuite is a local, full-featured companion to the hosted Forecast Dashboard
Shiny app. This vignette covers two ways to use it: the bundled app (`run_app()`), for
interactive, no-code forecasting, and the underlying model registry, for scripting
forecasts directly in R.

## The bundled app

```r
library(forecastsuite)
run_app()
```

`run_app()` launches a local Shiny app with four tabs, worked in order:

1. **Import Data** -- upload a file, pull a data frame from your global environment, a
   Google Sheet, a CSV URL, or pasted text; or pick the bundled demo dataset (five years
   of simulated daily clinic visits) to explore the app with no data of your own. An
   optional grouping column splits the finalized dataset into one series per distinct
   value (e.g. one per district), and a merge/relabel panel cleans up messy raw values
   ("female"/"Female"/"FEMALE") before they become separate groups.
2. **Holidays** -- Sundays, a fixed-date catalog, movable holidays from a file, and
   manual entries feed one compiled holiday list, with a consistency check against your
   data.
3. **Model** -- pick a model (Prophet, ARIMA, SARIMA, ETS, TBATS, NNETAR, Holt-Winters,
   or LSTM if `torch` is installed), Fit & Forecast, and browse the plot, metrics,
   rolling-origin cross-validation, and an **Analysis** panel: seasonal decomposition,
   anomaly detection, residual diagnostics, and (with a grouping column active)
   cross-group correlation. A "Show Code" panel prints the R equivalent of whatever you
   just ran, with a copy-to-clipboard button and a download.
4. **Model Guide** -- what each model is, its parameters, the evaluation metrics, and
   how the recommendation heuristic works.

Saving your setup (dataset, holidays, every Model-tab setting) to one file and reloading
it later is available from any tab, via "Save Project (.rds)" / "Load Project (.rds)".

## The registry API, without the app

Every model -- built-in or your own -- implements the same three-function contract:

```r
fit(train_df, date_agg, ...)         # -> model_obj
forecast(model_obj, h, ...)          # -> raw_forecast_obj
to_tibble(raw_forecast_obj, test_df) # -> tibble(ds, yhat[, yhat_lower, yhat_upper])
```

`train_df`/`test_df` are `ds`/`y` tibbles (date, value). `list_models()` lists what's
registered; `get_model(key)` returns one entry.

```{r}
library(forecastsuite)

vapply(list_models(available_only = TRUE), function(m) m$key, character(1))
```

### End to end on a synthetic series

```{r}
set.seed(1)
n <- 120
ds <- as.Date("2023-01-01") + 0:(n - 1)
y <- 100 + 15 * sin(2 * pi * seq_len(n) / 7) + seq_len(n) * 0.2 + rnorm(n, sd = 3)

train_df <- tibble::tibble(ds = ds[1:100], y = y[1:100])
test_df  <- tibble::tibble(ds = ds[101:120], y = y[101:120])

model <- get_model("arima")
fit_obj <- model$fit(train_df, date_agg = "day", auto = TRUE)
fc_raw  <- model$forecast(fit_obj, h = 20)
fc_tib  <- model$to_tibble(fc_raw, test_df)

head(fc_tib)
```

```{r}
safe_compute_metrics(fc_tib, test_df, label = "Test")
```

```{r}
plot_forecast_generic(fc_tib, train_df = train_df, subtitle = model$annotate(fit_obj))
```

### Analysis functions

The same functions behind the app's Analysis panel work standalone on any `ds`/`y`
tibble -- no fitted model needed for decomposition or anomaly detection, and no Shiny
session either.

```{r}
decomp <- decompose_series(train_df, date_agg = "day")
head(decomp)
```

```{r}
anomalies <- detect_anomalies(train_df, date_agg = "day", method = "iqr", threshold = 1.5)
anomalies[anomalies$is_anomaly, ]
```

```{r}
resid_diag <- compute_residual_diagnostics(fc_tib, test_df)
resid_diag[c("ljung_box_p", "shapiro_p", "n")]
```

`group_correlation_matrix()` takes a named list of `ds`/`y` tibbles (one per group, e.g.
`grouped_series()` from the app) and returns pairwise correlations:

```{r}
group_correlation_matrix(list(
  A = tibble::tibble(ds = ds, y = y),
  B = tibble::tibble(ds = ds, y = y * 0.5 + rnorm(n, sd = 1))
))
```

### Registering your own model

Adding a model is one `register_model()` call away from working everywhere the built-in
ones do (the app's Model tab, Compare Models, Cross-Validation, the Show Code panel):

```r
register_model(
  key = "naive_last",
  label = "Naive (last value)",
  fit = function(train_df, date_agg = "day", ...) utils::tail(train_df$y, 1),
  forecast = function(model_obj, h, ...) list(mean = rep(model_obj, h)),
  to_tibble = function(fc_obj, test_df) {
    tibble::tibble(ds = test_df$ds[seq_along(fc_obj$mean)], yhat = fc_obj$mean)
  }
)
```
